How to Analyze Build Logs --- A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-21 tse-paper/content.tex
grammarly

Despite its paramount role, the process of analyzing
build logs has so far received little attention, bringing us to our
first research question:
\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
1}}
What are the state-of-the-research techniques on build log analysis?
\end{simplebox}

In this article, as first step (\Cref{fig:overview}), we show that
less than half of the 61 works which rely on build log analysis
describe how they performed it.
Even fewer did so in detail or
made their implementation available.
Through the literature
survey we identify three prevalent strategies to
analyze build logs: Using 1) a set of
manually curated regular expressions,
similar to TravisTorrent~\cite{beller2017oops},
2) information
retrieval techniques,
and 3) a list of search keywords
to identify regions of interest in the build log.

@@

In step 3, we conduct an empirical study to gauge the performance of
these chunk retrieval techniques in
real-world projects under a variety of conditions.
Our study is based on the \emph{LogChunks} data
set~\cite{brandt2020logchunks} which encompasses build logs from
80 popular open source projects and a broad range of
build environments.
Finally, we develop guidelines (step 4) on when
one of the techniques might be preferential to its alternatives.

@@

To enable retracing of our research, we provide a detailed replication
package~\cite{brandt2020chunk-replication}.

@@

The field of analyzing build
logs is relatively young; We have seen scattered
research efforts on build log analysis.
However,the processing of build logs was often
a side-part of the article.
To the best of our
knowledge, there has not been an attempt to systematically overview
and survey approaches for processing build logs.
We conducted a systematic mapping study to determine the
state-of-research on build log analysis following the guidelines
by Petersen et al.~\cite{petersen2008systematic,petersen2015guidelines}.

@@

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A related field of build log analysis is the processing of system log
files produced during runtime.
A main difference between build logs
and system logs is that system logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} shows
example lines from a system log.

@@

Our Google Scholar search for variations of ``build log''
returns more than 2.2 million results (2020-03-15).
We refined our search criterions to exclude obvious
non-related works.
For example, we decided to exclude systems
or event logs (see \Cref{sec:system-log-analysis}).
This left us with 1,552 search results
(2020-03-15).
The full search queries are documented in our replication
package~\cite{brandt2020chunk-replication}.

@@

With these 256 remaining articles, we followed a backward snowball
sampling for related work.
If one one of the articles referenced a new article in the context
of build log analysis we added the new to our literature set.
This added five papers (eight before
duplicate elimination), coming to a total of 261
articles.

Subsequently, we performed a second, finer filter phase by downloading the
full articles, reading their abstracts, and searching the full text
for the occurrence of ``log''.
If these papers showed traces of
working with build logs, we included them in our survey.
In total, we
were left with 87 articles after the two filtering phases.

@@

Relevant for our survey were all articles that
use logs from software builds as input to
some kind of process.
This excluded 15 articles from our study.
For several cases, multiple articles were extensions of others.
Of these, we only considered the earliest article which reported on
the processing of build logs.
This excluded another 11 articles.
In sum, we further extracted data from 61 articles.

@@

\begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ1:} What are the state-of-the-research
techniques on build log analysis?}
\begin{itemize}[leftmargin=2cm]
  \item[\textbf{RQ1.1:}] Which information is targeted?
  \item[\textbf{RQ1.2:}] Which techniques are used?
  \item[\textbf{RQ1.3:}] Which build logs are analyzed?
\end{itemize}
\end{simplebox}

@@

In most cases (51\%), a chunk was retrieved from the log.
Build logs were scanned for compilation
errors~\cite{clemencic2014new} or
the duration of tasks within the build~\cite{zhang2016android}.
36\% of the articles classified the build logs into categories.
For instance distinguishing failures caused by developers from
failures caused
by infrastructure errors~\cite{lindqvist2019detection} or
whether projects use automated static analysis
tools~\cite{kavaler2019tool}.
Other types of information (e.g.\ counting
lines, summarization) were only retrieved
in one case each.

@@

The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs''.
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
Another 8\% employed machine learning techniques such as natural language
processing or information retrieval techniques.
To give examples,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They start with manual categorization of the build logs and select
search strings that identify their targeted categories.
Based on the search strings they create a script which automatically
classifies the remaining logs.

25\% of the articles claimed their implementation is available,
however, we
saw very low reuse of build log analysis tools within our survey.
Five articles employed the tools used to create
\emph{TravisTorrent}~\cite{beller2017travistorrent,beller2017oops,
orellana2017differences,zhao2018comparing} or
an extended version of them~\cite{rott2019empirische,
shi2018evaluating},
two used the
\emph{Maven Log Analyzer}~\cite{macho2018automatically,gallaba2018noise},
another two the
\emph{MAKAO}~\cite{wen2018blimp,adams2007design,adams2007makao} tool.

@@

Only 7\% of the proposed methods claimed to be language-agnostic, while
several described they are covering multiple source build tools.
The majority of articles (42\%) targeted logs from a source which was not
targeted by any second article.

The main platform that build logs were collected from is Travis
CI~\cite{travisci2019webpage},
which was used in 38\% of cases.
18\% of the articles processed logs from TravisTorrent and 8\% included
build logs from industrial projects.

@@

31\% of the surveyed articles claimed to have published their data to
enable further research.
However we found no clear reuse of build log data sets with the
exception of TravisTorrent~\cite{beller2017travistorrent}, which was
analyzed by Vassallo et al.~\cite{vassallo2017a-tale},
Ghaleb et al.~\cite{ghaleb2019studying} and several other studies.

@@

Several articles noted the variety in log formats for different build
environments, which requires customizing build log techniques for
every supported build environment.
\Cref{fig:litsur:log_producer} shows that few build environments are
targeted by many studies and tools (A).
The majority of build environments was however only targeted by one
article each (B).
From this we follow that, while custom parsers are beneficial for the
small group of environments targeted by many (A), all and especially
the rarer investigated environments (B) would benefit from a generic
solution.
% We implement build environment agnostic build log analysis
% techniques, which are customized to a specific project by
% the training examples which the user provides.
% In addition, the \emph{LogChunks}~\cite{brandt2020logchunks}
% data set we chose for our evaluation covers a broad range of build
% environments.
% % Further, the training examples also specify the log chunk targeted
% by the
% % retrieval, which enable our chunk retrieval techniques to retrieve
% % a broad range of information from build logs.

@@

For this reason, we are focussing on techniques that run automatic
after being configured through examples or keywords.
Examples consist of logs and corresponding chunks.
The examples and keywords are provided by the user and can simply be
exchanged whenever new cases are introduced.

We propose the notion of chunk retrieval techniques which retrieve
a substring, that describes a specific information, from a build log.
Generally these can also be used to classify build logs and therefore
cover the great majority of the type of log analysis we saw within
our survey.

@@

Our evaluation of the three chunk retrieval techniques is done on the
\emph{LogChunks} data set, because it encompasses a broad range of
build environments which enables us to measure the generic
applicability of our techniques.
The targeted chunks are describing error messages or
``the reason the build failed'', which was also the information
most targeted by the articles in our survey.

@@

This section introduces key concepts of the three chunk retrieval
techniques we study in this article, namely program synthesis by
example (PBE), common text similarity (CTS), and keyword search (KWS).
In addition we present our comparison baseline of random line
retrieval (RLR), illustrate the techniques on a concrete example
of log chunks and sketch out implementation.

@@

With the term \textit{configuration}, we abstract over the training
and parametrization that different techniques require in different
forms.
A configuration can be explicitly stated or implicitly derived
by learning through provided examples.
It is therefore a manual
specification of which information the chunk retrieval should target.
It also supplies the necessary information for the technique to
identify the targeted information chunk in a build log.
Each chunk
retrieval technique has a specific \textit{granularity}, i.e.,\ the
smallest piece of text it can return (e.g., a line, or a word).
The
granularity might be adjustable by configuration.
\emph{Running a
chunk retrieval technique} means to execute a fully configured
technique to consumes as input a build log in plain text format and to
produce an array as output.
The array consists of substrings of the
build log text.

\subsection{Program Synthesis by Example (PBE)}
\label{sec:expl-pbe}
The concept of \emph{Programming by Example} aims to capture the
intent of the user through examples which they provide.
Leveraging this approach, we implemented a chunk retrieval technique
for build logs.
Our implementation is based and highly influenced by the
\emph{Microsoft PROSE library}~\cite{prose2019webpage}.
The PROSE library is powered by the generic program synthesis framework
\emph{FlashMeta}~\cite{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\cite{le2014flashextract:}.
Both enable us to synthesize regular expression programs
consistent~\cite{mitchell1982generalization} with a set of in-
and output examples given by the user.
The usage of example enables the user to configure a chunk
retrieval without understanding the whole structure of
the build log.

@@

\subsection{Common Text Similarity (CTS)}
\label{sec:expl-ts}
Text Similarity approaches are widely used to filter unstructured
textual software artifacts~\cite{runeson2007detection,
marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
We investigate a chunk retrieval technique based on the common
Vector Space Model~\cite{schutze2008introduction}.
The aim is to retrieve those lines from a build log which are most
similar to examples given by the user.

\subsubsection{Configuration}
To configure chunk retrieval though text similarity we chose to use
the same concept of examples as for PBE
% TODO Moritz: your question was: More details on this.
% How do you
% 'apply' VSM? On what?
% all the following sentences in this subsubsection describe our way
% of determining text similarity which as far as I understand and
% Annibale said is the VSM.
% word tokens, tf-idf, similarity measure
% I added a "as follows", think that is enough for explanation?
and apply the Vector Space Model~\cite{schutze2008introduction}
as follows:
The lines of the output strings of the training examples define our
search query.
The algorithm splits the search query into single lines and
identifies tokens, in our case words.
Then we build a
document-term-frequency matrix over the lines from the search query
and prune very often or very rarely appearing words.
Finally, the
algorithm applies TF-IDF to the matrix, a best practice for natural
language queries~\cite{lee1997document}.

@@

\subsection{Keyword Search (KWS)}
\label{sec:expl-skws}
When developers scavenge for a specific piece of information within a
large amount of unstructured information, a first ad-hoc approach they
use is to search for related keywords.
Indeed, this was one of the
most common approaches we took when searching for the reason a build
failed while creating the \emph{LogChunks} data
set~\cite{brandt2020logchunks}.

@@

\subsection{Tool Implementation}
For our comparison study we implemented each of the chunk retrieval
techniques and a unifying interface.
The unified interface is implemented in Ruby and
calls the separate technique implementations over the command line.
The implementation of PBE in C\# is based on the Microsoft PROSE
library~\cite{prose2019webpage}.
We implemented CTS, KWS and RLR using
R and the text2vec library~\cite{text2vec2019webpage}.
They can be found in our replication
package~\cite{brandt2020chunk-replication} together which have
our techniques installed and ready to use.

@@

To investigate when our chunk retrieval techniques are suited to
analyze build logs (\textbf{RQ2}), we perform an empirical comparison
study on the \emph{LogChunks} data set~\cite{brandt2020logchunks}.
In particular we are interested in how many examples a technique
should be trained with (\textbf{RQ2.1}),
how similar the structural representation of the targeted log chunks
has to be (\textbf{RQ2.2}) and how reliable the quality of the
produced outputs is (\textbf{RQ2.3}).

@@

In the following we present the data set we created for this study,
our study process (presented in \Cref{fig:study}) and the metrics
we measure to answer our research questions.

@@

\subsection{LogChunks}
For this study we created the previously published data set
\emph{LogChunks}.
It encompasses 797 build logs from Travis CI,
stemming from a broad range of 80 GitHub repositories and 29
programming languages.
For each file, the authors manually ``labeled
the log part (chunk) describing why the build
failed''~\cite{brandt2020logchunks}.
We included keywords, which we
would use to search for the selected chunk within the log.
In
addition, we categorized the log chunks according to their format
within the log.
If the chunks were surrounded by the same markings
within the log we assigned them the \emph{same structural category},
as described in \Cref{sec:crt-example}.

\subsection{Study Process}
We begin by selecting all examples in \emph{LogChunks}
that correspond to one GitHub repository---and thus the
same build environment (1).
One to five examples make up the training set (2).
We vary its size to measure how many examples are needed to
confidently configure a technique.
The training set is so small, because the examples have to be
manually created by a user.
A larger training set would oppose our goal of providing techniques
that can be configured with little effort
(see \Cref{sec:lit-sur:discussion}).
The example chronologically after the training set makes up the
test set (3).
In this way, we train on examples from past build logs and test on
more recent logs.
We chose a test set size of 1, as users of our techniques
should receive reliable results on every build log they analyze.
In the next step, we configure a chunk retrieval technique with
the examples from the training set (4) and apply the technique to the
training set (5), which produces the retrieval output (6).
Based on this, we collect the following measurements (7):
the size of training set (\textbf{RQ2.1}),
the number of structural categories in the training
set (\textbf{RQ2.2}),
the lines of the retrieval output ($\mathit{RetrievedLines}$),
and our oracle: the output defined in the test example
($\mathit{DesiredLines}$).
From these we calculate accuracy metrics (\textbf{RQ2.3}):

@@

\section{Results}
This section presents the separate results for each chunk retrieval
technique.
Afterwards we compare the three techniques with each other and RLR as
baseline.

@@

In our evaluation, we compare against a baseline of randomly
picking lines from the build log.
Its results follow intuitive
expectations:
The precision ranges between 8\% and 5\%, the recall between 6\% and
8\%.
The size of the training set has no visible impact on either.
An larger number of structural categories within the training and
test sets decreased the precision of RLR from 7\% to 0\%, the recall
increases from 6\%, for one structural category present, to 11\%, for
three structural categories present and also drops to 0\% for more
structural categories.
Graphs of these results of the chunk retrieval with RLR
are included in our replication
package~\cite{brandt2020chunk-replication}.

@@

For the interpretation of our study results we look at each chunk
retrieval technique separately and give recommendations for
which types of input build
logs, available training examples and consumption of the retrieved
output each technique is suited.
Following that, we discuss
which of these criteria should influence the decision to use a certain
chunk retrieval technique most.
Based on our empirical comparison, we present a
decision tree between the three techniques we investigated.

@@

\subsubsection{Keyword Search (KWS)}
\noindent
\textbf{Configuration and Input}
KWS has a higher recall than the two other techniques for multiple
structural categories present in the training and test examples.
This
makes KWS a good technique if there is little prior knowledge on how
the targeted log chunk is represented in the build log.
For the
example of extracting the reason the build failed, KWS is best suited
if a build can fail in various steps logged by different tools and no
pre-categorization of where the build failed is available.

@@

If the information chunks are from multiple structural categories
and recall is more important than precision we recommend
to use KWS\@.
If precision is more important than recall to the user we
recommend CTS\@.
We also recommend CTS when the log chunks are
from one structural category, when the user does not require a high
confidence in the precision or recall of the outcome and when one
would rather have output with low recall instead of no output at all.
When the representations are from one structural category and the user
wishes a high confidence in the correctness of the output or prefers
no output over output with low recall, we recommend PBE\@.

As a final recommendation, one could create a ``super analyzer'' by
combining the different build log analysis techniques studied in this
paper.
Such a super analyzer would likely always first employ PBE
(because of its high accuracy), followed by a combination of CTS and
KWS\@.
Other than initial setup and training time, there should be no
downside to this approach if it is implemented in a transparent way to
the underlying technique: the output of the super analyzer could
include from which sub-tool it originated and thus facilitate
automatic on-ward processing or interpretation of the result.

@@

\textbf{Refinement of Retrieval Quality for each Technique} We
investigated basic configurations of existing techniques applied to
chunk retrieval from build logs.
In a next step, each of these
techniques could be refined to better approach the domain of build
logs.
The \emph{LogChunks} data set and our study results act as a
baseline to benchmark such technique improvements.
We propose the
following improvements:
\begin{itemize}[leftmargin=0.4cm]
  \item \textbf{Custom Ranking and Tokens for PBE} The program
  synthesis through PROSE ranks possible programs according to
  what the user most likely intended.
  One could adapt the ranking
  rules provided by the FlashExtract DSL to fit common build log
  chunk retrieval tasks.
  FlashExtract includes special tokens when
  enumerating possible regular expressions.
  One could extend these
  with tokens found in build logs, such as ``-'',``='',``ERROR''
  or ``[OK''.
  \item \textbf{Meta-Parameter Optimization for CTS} Information
  retrieval techniques have various meta-parameters which can be
  optimized for the specific use
  case~\cite{panichella2016parameterizing}.
  We propose to further
  investigate improvements in preprocessing of the log text, in
  tokenization of the log lines into terms and in stop words
  lists.
\end{itemize}

>>>>

Despite its paramount role, the process of analyzing
build logs has so far received little attention, bringing us to our
first research question:
\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
1}}
What are the state-of-the-research techniques for build log analysis?
\end{simplebox}

In this article, as the first step (\Cref{fig:overview}), we show that
less than half of the 61 works which rely on build log analysis
describe how they performed it.
Even fewer did so in detail or
made their implementation available.
Through the literature
survey we identify three prevalent strategies to
analyze build logs: Using 1) a set of
manually curated regular expressions,
similar to TravisTorrent~\cite{beller2017oops},
2) information
retrieval techniques,
and 3) a list of search keywords
to identify regions of interest in the build log.

@@

In step 3, we conduct an empirical study to gauge the performance of
these chunk retrieval techniques in
real-world projects under a variety of conditions.
Our study is based on the \emph{LogChunks} data
set~\cite{brandt2020logchunks} which encompasses build logs from
80 popular open-source projects and a broad range of
build environments.
Finally, we develop guidelines (step 4) on when
one of the techniques might be preferential to its alternatives.

@@

To enable the retracing of our research, we provide a detailed replication
package~\cite{brandt2020chunk-replication}.

@@

The field of analyzing build
logs is relatively young; We have seen scattered
research efforts on build log analysis.
However, the processing of build logs was often
a side-part of the article.
To the best of our
knowledge, there has not been an attempt to systematically overview
and survey approaches for processing build logs.
We conducted a systematic mapping study to determine the
state-of-research on build log analysis following the guidelines
by Petersen et al.~\cite{petersen2008systematic,petersen2015guidelines}.

@@

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A related field of build log analysis is the processing of system log
files produced during runtime.
The main difference between build logs
and system logs is that system logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} shows
example lines from a system log.

@@

Our Google Scholar search for variations of ``build log''
returns more than 2.2 million results (2020-03-15).
We refined our search criteria to exclude obvious
non-related works.
For example, we decided to exclude systems
or event logs (see \Cref{sec:system-log-analysis}).
This left us with 1,552 search results
(2020-03-15).
The full search queries are documented in our replication
package~\cite{brandt2020chunk-replication}.

@@

With these 256 remaining articles, we followed a backward snowball
sampling for related work.
If one of the articles referenced a new article in the context
of build log analysis we added the new to our literature set.
This added five papers (eight before
duplicate elimination), coming to a total of 261
articles.

Subsequently, we performed a second, finer filter phase by downloading the
full articles, reading their abstracts and searching the full text
for the occurrence of ``log''.
If these papers showed traces of
working with build logs, we included them in our survey.
In total, we
were left with 87 articles after the two filtering phases.

@@

Relevant for our survey were all articles that
use logs from software builds as input to
some kind of process.
This excluded 15 articles from our study.
In several cases, multiple articles were extensions of others.
Of these, we only considered the earliest article which reported on
the processing of build logs.
This excluded another 11 articles.
In sum, we further extracted data from 61 articles.

@@

\begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ1:} What are the state-of-the-research
techniques for build log analysis?}
\begin{itemize}[leftmargin=2cm]
  \item[\textbf{RQ1.1:}] Which information is targeted?
  \item[\textbf{RQ1.2:}] Which techniques are used?
  \item[\textbf{RQ1.3:}] Which build logs are analyzed?
\end{itemize}
\end{simplebox}

@@

In most cases (51\%), a chunk was retrieved from the log.
Build logs were scanned for compilation
errors~\cite{clemencic2014new} or
the duration of tasks within the build~\cite{zhang2016android}.
36\% of the articles classified the build logs into categories.
For instance, distinguishing failures caused by developers from
failures caused
by infrastructure errors~\cite{lindqvist2019detection} or
whether projects use automated static analysis
tools~\cite{kavaler2019tool}.
Other types of information (e.g.\ counting
lines, summarization) were only retrieved
in one case each.

@@

The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs''.
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
Another 8\% employed machine learning techniques such as natural language
processing or information retrieval techniques.
To give examples,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They start with manual categorization of the build logs and select
search strings that identify their targeted categories.
Based on the search strings they create a script that automatically
classifies the remaining logs.

25\% of the articles claimed their implementation is available,
however, we
saw a very low reuse of build log analysis tools within our survey.
Five articles employed the tools used to create
\emph{TravisTorrent}~\cite{beller2017travistorrent,beller2017oops,
orellana2017differences,zhao2018comparing} or
an extended version of them~\cite{rott2019empirische,
shi2018evaluating},
two used the
\emph{Maven Log Analyzer}~\cite{macho2018automatically,gallaba2018noise},
another two the
\emph{MAKAO}~\cite{wen2018blimp,adams2007design,adams2007makao} tool.

@@

Only 7\% of the proposed methods claimed to be language-agnostic, while
several described they are covering multiple source build tools.
The majority of articles (42\%) targeted logs from a source that was not
targeted by any second article.

The main platform that build logs were collected from is Travis
CI~\cite{travisci2019webpage},
which was used in 38\% cases.
18\% of the articles processed logs from TravisTorrent and 8\% included
build logs from industrial projects.

@@

31\% of the surveyed articles claimed to have published their data to
enable further research.
However, we found no clear reuse of build log data sets with the
exception of TravisTorrent~\cite{beller2017travistorrent}, which was
analyzed by Vassallo et al.~\cite{vassallo2017a-tale},
Ghaleb et al.~\cite{ghaleb2019studying} and several other studies.

@@

Several articles noted the variety in log formats for different build
environments, which requires customizing build log techniques for
every supported build environment.
\Cref{fig:litsur:log_producer} shows that few build environments are
targeted by many studies and tools (A).
The majority of build environments was however only targeted by one
article each (B).
From this, we follow that, while custom parsers are beneficial for the
small group of environments targeted by many (A), all and especially
the rarer investigated environments (B) would benefit from a generic
solution.
% We implement build environment agnostic build log analysis
% techniques, which are customized to a specific project by
% the training examples which the user provides.
% In addition, the \emph{LogChunks}~\cite{brandt2020logchunks}
% data set we chose for our evaluation covers a broad range of build
% environments.
% % Further, the training examples also specify the log chunk targeted
% by the
% % retrieval, which enable our chunk retrieval techniques to retrieve
% % a broad range of information from build logs.

@@

For this reason, we are focussing on techniques that run automatically
after being configured through examples or keywords.
Examples consist of logs and corresponding chunks.
The examples and keywords are provided by the user and can simply be
exchanged whenever new cases are introduced.

We propose the notion of chunk retrieval techniques which retrieve
a substring, that describes a specific information, from a build log.
Generally, these can also be used to classify build logs and therefore
cover the great majority of the type of log analysis we saw within
our survey.

@@

Our evaluation of the three chunk retrieval techniques is done on the
\emph{LogChunks} data set because it encompasses a broad range of
build environments which enables us to measure the generic
applicability of our techniques.
The targeted chunks are describing error messages or
``the reason the build failed'', which was also the information
most targeted by the articles in our survey.

@@

This section introduces key concepts of the three chunk retrieval
techniques we study in this article, namely program synthesis by
example (PBE), common text similarity (CTS), and keyword search (KWS).
In addition, we present our comparison baseline of random line
retrieval (RLR), illustrate the techniques on a concrete example
of log chunks and sketch out implementation.

@@

With the term \textit{configuration}, we abstract over the training
and parametrization that different techniques require in different
forms.
A configuration can be explicitly stated or implicitly derived
by learning through provided examples.
It is therefore a manual
specification of which information the chunk retrieval should target.
It also supplies the necessary information for the technique to
identify the targeted information chunk in a build log.
Each chunk
retrieval technique has a specific \textit{granularity}, i.e.,\ the
smallest piece of text it can return (e.g., a line, or a word).
The
granularity might be adjustable by configuration.
\emph{Running a
chunk retrieval technique} means executing a fully configured
technique to consumes as input a build log in plain text format and to
produce an array as output.
The array consists of substrings of the
build log text.

\subsection{Program Synthesis by Example (PBE)}
The concept of \emph{Programming by Example} aims to capture the
intent of the user through examples which they provide.
Leveraging this approach, we implemented a chunk retrieval technique
for build logs.
Our implementation is based and highly influenced by the
\emph{Microsoft PROSE library}~\cite{prose2019webpage}.
The PROSE library is powered by the generic program synthesis framework
\emph{FlashMeta}~\cite{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\cite{le2014flashextract:}.
Both enable us to synthesize regular expression programs
consistent~\cite{mitchell1982generalization} with a set of in-
and output examples given by the user.
The usage of example enables the user to configure a chunk
retrieval without understanding the whole structure of
the build log.

@@

\subsection{Common Text Similarity (CTS)}
Text Similarity approaches are widely used to filter unstructured
textual software artifacts~\cite{runeson2007detection,
marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
We investigate a chunk retrieval technique based on the common
Vector Space Model~\cite{schutze2008introduction}.
The aim is to retrieve those lines from a build log which are most
similar to examples given by the user.

\subsubsection{Configuration}
To configure chunk retrieval through text similarity we chose to use
the same concept of examples as for PBE
% TODO Moritz: your question was: More details on this.
% How do you
% 'apply' VSM? On what?
% all the following sentences in this subsubsection describe our way
% of determining text similarity which as far as I understand and
% Annibale said is the VSM.
% word tokens, tf-idf, similarity measure
% I added a "as follows", think that is enough for explanation?
and apply the Vector Space Model~\cite{schutze2008introduction}
as follows:
The lines of the output strings of the training examples define our
search query.
The algorithm splits the search query into single lines and
identifies tokens, in our case words.
Then we build a
document-term-frequency matrix over the lines from the search query
and prune very often or very rarely appearing words.
Finally, the
algorithm applies TF-IDF to the matrix, a best practice for natural
language queries~\cite{lee1997document}.

@@

\subsection{Keyword Search (KWS)}
When developers scavenge for a specific piece of information within a
large amount of unstructured information, a first ad-hoc approach they
use is to search for related keywords.
Indeed, this was one of the
most common approaches we took when searching for the reason a build
failed while creating the \emph{LogChunks} data
set~\cite{brandt2020logchunks}.

@@

\subsection{Tool Implementation}
For our comparison study, we implemented each of the chunk retrieval
techniques and a unifying interface.
The unified interface is implemented in Ruby and
calls the separate technique implementations over the command line.
The implementation of PBE in C\# is based on the Microsoft PROSE
library~\cite{prose2019webpage}.
We implemented CTS, KWS and RLR using
R and the text2vec library~\cite{text2vec2019webpage}.
They can be found in our replication
package~\cite{brandt2020chunk-replication} together
with Docker images which have
our techniques installed and ready to use.

@@

To investigate when our chunk retrieval techniques are suited to
analyze build logs (\textbf{RQ2}), we perform an empirical comparison
study on the \emph{LogChunks} data set~\cite{brandt2020logchunks}.
In particular, we are interested in how many examples a technique
should be trained with (\textbf{RQ2.1}),
how similar the structural representation of the targeted log chunks
has to be (\textbf{RQ2.2}) and how reliable the quality of the
produced outputs is (\textbf{RQ2.3}).

@@

In the following, we present the data set we created for this study,
our study process (presented in \Cref{fig:study}) and the metrics
we measure to answer our research questions.

@@

\subsection{LogChunks}
For this study, we created the previously published data set
\emph{LogChunks}.
It encompasses 797 build logs from Travis CI,
stemming from a broad range of 80 GitHub repositories and 29
programming languages.
For each file, the authors manually ``labeled
the log part (chunk) describing why the build
failed''~\cite{brandt2020logchunks}.
We included keywords, which we
would use to search for the selected chunk within the log.
In
addition, we categorized the log chunks according to their format
within the log.
If the chunks were surrounded by the same markings
within the log, we assigned them the \emph{same structural category},
as described in \Cref{sec:crt-example}.

\subsection{Study Process}
We begin by selecting all examples in \emph{LogChunks}
that correspond to one GitHub repository---and thus the
same build environment (1).
One to five examples make up the training set (2).
We vary its size to measure how many examples are needed to
confidently configure a technique.
The training set is so small because the examples have to be
manually created by a user.
A larger training set would oppose our goal of providing techniques
that can be configured with little effort
(see \Cref{sec:lit-sur:discussion}).
The example chronologically after the training set makes up the
test set (3).
In this way, we train on examples from past build logs and test on
more recent logs.
We chose a test set size of 1, as users of our techniques
should receive reliable results on every build log they analyze.
In the next step, we configure a chunk retrieval technique with
the examples from the training set (4) and apply the technique to the
training set (5), which produces the retrieval output (6).
Based on this, we collect the following measurements (7):
the size of the training set (\textbf{RQ2.1}),
the number of structural categories in the training
set (\textbf{RQ2.2}),
the lines of the retrieval output ($\mathit{RetrievedLines}$),
and our oracle: the output defined in the test example
($\mathit{DesiredLines}$).
From these we calculate accuracy metrics (\textbf{RQ2.3}):

@@

\section{Results}
This section presents the separate results for each chunk retrieval
technique.
Afterwards, we compare the three techniques with each other and RLR as
baseline.

@@

In our evaluation, we compare against a baseline of randomly
picking lines from the build log.
Its results follow intuitive
expectations:
The precision ranges between 8\% and 5\%, the recall between 6\% and
8\%.
The size of the training set has no visible impact on either.
A larger number of structural categories within the training and
test sets decreased the precision of RLR from 7\% to 0\%, the recall
increases from 6\%, for one structural category present, to 11\%, for
three structural categories present and also drops to 0\% for more
structural categories.
Graphs of these results of the chunk retrieval with RLR
are included in our replication
package~\cite{brandt2020chunk-replication}.

@@

For the interpretation of our study results, we look at each chunk
retrieval technique separately and give recommendations for
which types of input build
logs, available training examples and consumption of the retrieved
output each technique is suited.
Following that, we discuss
which of these criteria should influence the decision to use a certain
chunk retrieval technique most.
Based on our empirical comparison, we present a
decision tree between the three techniques we investigated.

@@

\subsubsection{Keyword Search (KWS)}
\noindent
\textbf{Configuration and Input}
KWS has a higher recall than the two other techniques for multiple
structural categories present in the training and test examples.
This
makes KWS a good technique if there is little prior knowledge of how
the targeted log chunk is represented in the build log.
For the
example of extracting the reason the build failed, KWS is best suited
if a build can fail in various steps logged by different tools and no
pre-categorization of where the build failed is available.

@@

If the information chunks are from multiple structural categories
and recall is more important than precision we recommend
to use KWS\@.
If precision is more important than recall to the user we
recommend CTS\@.
We also recommend CTS when the log chunks are
from one structural category, when the user does not require high
confidence in the precision or recall of the outcome and when one
would rather have output with low recall instead of no output at all.
When the representations are from one structural category and the user
wishes high confidence in the correctness of the output or prefers
no output over output with low recall, we recommend PBE\@.

As a final recommendation, one could create a ``super analyzer'' by
combining the different build log analysis techniques studied in this
paper.
Such a super analyzer would likely always first employ PBE
(because of its high accuracy), followed by a combination of CTS and
KWS\@.
Other than initial setup and training time, there is no
downside to this approach if it is implemented in a transparent way to
the underlying technique: the output of the super analyzer could
include from which sub-tool it originated and thus facilitate
automatic on-ward processing or interpretation of the result.

@@

\textbf{Refinement of Retrieval Quality for each Technique} We
investigated basic configurations of existing techniques applied to
chunk retrieval from build logs.
In a next step, each of these
techniques could be refined to better approach the domain of build
logs.
The \emph{LogChunks} data set and our study results act as a
baseline to benchmark such technique improvements.
We propose the
following improvements:
\begin{itemize}[leftmargin=0.4cm]
  \item \textbf{Custom Ranking and Tokens for PBE} The program
  synthesis through PROSE ranks possible programs according to
  what the user most likely intended.
  One could adapt the ranking
  rules provided by the FlashExtract DSL to fit common build log
  chunk retrieval tasks.
  FlashExtract includes special tokens when
  enumerating possible regular expressions.
  One could extend these
  with tokens found in build logs, such as ``-'',
	``='', ``ERROR''
  or ``[OK''.
  \item \textbf{Meta-Parameter Optimization for CTS} Information
  retrieval techniques have various meta-parameters which can be
  optimized for the specific use
  case~\cite{panichella2016parameterizing}.
  We propose to further
  investigate improvements in preprocessing of the log text, in
  tokenization of the log lines into terms and in stop words
  lists.
\end{itemize}
